A Type System for String Sanitation\\Implemented Inside a Python

### 2014-08-01 sanitation-psp14/sanitation-psp14.tex
Added PSP changes.

\documentclass[10pt,preprint]{sigplanconf}
\newcommand{\lamAce}{\lambda_{\text{Ace}}}
% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.
\usepackage{afterpage}
\usepackage{listings}
\usepackage[usenames,dvipsnames]{color}
\usepackage{mathpartir}
\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{amsthm}
\usepackage{ stmaryrd }
\usepackage{verbatimbox}
\input{../att-icfp14/macros-atlam}
\usepackage{alltt}
\renewcommand{\ttdefault}{txtt}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\definecolor{light-gray}{gray}{0.5}
\usepackage{listings}
\usepackage{wasysym}
    \makeatletter
\usepackage{enumitem}
\usepackage{enumerate}

@@

% Source language terms.
\newcommand{\sisubst}[3]{{\sf rreplace}[#1](#2,#3)}
\newcommand{\rssreplace}[3]{\sisubst{#1}{#2}{#2}} % TODO-nrf fix this.

\newcommand{\coerce}[2]{ {\sf rcoerce}[#1](#2)}
\newcommand{\sistr}[1]{{\sf rstr}[#1]}
\newcommand{\strin}[1]{\sistr{#1}}
\newcommand{\rsconcat}[2]{{\sf rconcat}(#1,#2)}

@@

% target language terms.
\newcommand{\tsubst}[3]{{\sf replace}(#1,#2,#3)} % TODO-nrf rename the commands.
\newcommand{\tcheck}[2]{{\sf check}(#1, #2)}
\renewcommand{\tstr}[1]{{{\sf str}[#1]}}
\newcommand{\tconcat}[2]{{\sf concat}(#1,#2)}
\newcommand{\regext}[1]{ {\sf rx}[#1] } % TODO-nrf remove
\newcommand{\rx}[1]{ {\sf rx}[#1] }

@@

\conferenceinfo{-}{-} 
\copyrightyear{-} 
\copyrightdata{[to be supplied]}

@@

%\authorinfo{~}{~}{~}
\authorinfo{Nathan Fulton\and Cyrus Omar\and Jonathan Aldrich}
           {School of Computer Science\\
            Carnegie Mellon University}
           {\{nathanfu, comar, aldrich\}@cs.cmu.edu}

@@

@@

@@

\afterpage{

@@

\inferrule[P-T-Check]
{\tctx \vdash P_1 : \regex \\ \tctx \vdash P_2 : \str \\ }
{\tctx \vdash \tcheck{P_1}{P_2} : \str}

\end{mathpar}
\caption{Typing rules for our fragment of $\lambdap$.
The typing context $\tctx$ is standard.}
\label{fig:etyping}
\end{figure}

@@

\clearpage
} % End \afterpage

%
% Begin translation
%

\begin{figure}[t]

\end{mathpar}
\caption{Translation from source terms (S) to target terms (P).
The translation is type-directed in the Tr-Coerce cases.}
\end{figure}

Note that we never insert an internal check where the type of the string implies
that a check must succeed. Furthermore, expensive calculations (such as language
inclusion or ${\tt dsubst}$, all occur at compile time.
Since sanitation problems are generally expressible with smalle, simple expressions,
we believe that the compile-time overhead is not significant enough to prohibiour target usecase. Informal experimentation with our implementation seems to
support this assertion.

@@

\begin{thm}
  Let $S$ be a term in the source language. If $\sctx \vdash S : \psi$ 
  then $S \sreduces S'$ and $S' : \psi$.
\end{thm}
\begin{proof}
  TODO-nrf

By a straightforward induction on the derivation of $\sctx \vdash S : \psi$.
Note that preservation depends upon the substitution correspondence theorem
for the T-T-Subst case and the closure lemmas for the T-T-Concat case.
\end{proof}

\begin{thm}
  Let $P$ be a term in the target language. If $\tctx \vdash P : \theta$ 
  then $P \sreduces P'$ and $P' : \theta$.
\end{thm}
\begin{proof}
TODO-nrf

  By a straightforward induction on the derivation of $\tctx \vdash P : \theta$.
Note that the well-formedness of regular expressions indexing ${\tt rx}$ types
depends upon the closure lemmas.
\end{proof}

>>>>

\documentclass{acm_proc_article-sp}
\newcommand{\lamAce}{\lambda_{\text{Ace}}}
% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.
\usepackage{afterpage}
\usepackage{listings}
\usepackage[usenames,dvipsnames]{color}
\usepackage{mathpartir}
\usepackage{amsmath}
\usepackage{amssymb}

% Hack. Something's wrong with PLAS paper when using the ACM Proc docclass
\let\proof\relax
\let\endproof\relax
\usepackage{amsthm}

\usepackage{ stmaryrd }
\usepackage{verbatimbox}
\input{../att-icfp14/macros-atlam}
\usepackage{alltt}
\renewcommand{\ttdefault}{txtt}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\definecolor{light-gray}{gray}{0.5}
\usepackage{listings}
\usepackage{wasysym}
    \makeatletter
\usepackage{enumitem}
\usepackage{enumerate}

@@

% Source language terms.
\newcommand{\sisubst}[3]{{\sf rreplace}[#1](#2,#3)} \newcommand{\rreplace}[3]{{\sf rreplace}[#1](#2,#3)} % lots of legacy naming around these parts...

\newcommand{\rssreplace}[3]{\sisubst{#1}{#2}{#2}} % TODO-nrf fix this.

\newcommand{\coerce}[2]{ {\sf rcoerce}[#1](#2)}
\newcommand{\sistr}[1]{{\sf rstr}[#1]}   \newcommand{\rstr}[1]{{\sf rstr}[#1]} % Lots of legacy naming around these parts...

\newcommand{\strin}[1]{\sistr{#1}}
\newcommand{\rsconcat}[2]{{\sf rconcat}(#1,#2)}

@@

% target language terms.
\newcommand{\tsubst}[3]{{\sf replace}(#1,#2,#3)} \newcommand{\metareplace}[3]{{\sf replace}(#1,#2,#3)} % TODO-nrf rename the commands. Lots of legacy naming around these parts...

\newcommand{\tcheck}[2]{{\sf check}(#1, #2)}
\renewcommand{\tstr}[1]{{{\sf str}[#1]}}
\newcommand{\tconcat}[2]{{\sf concat}(#1,#2)}
\newcommand{\regext}[1]{ {\sf rx}[#1] } % TODO-nrf remove
\newcommand{\rx}[1]{ {\sf rx}[#1] }

@@

%dep_docclass\conferenceinfo{-}{-} 
%dep_docclass\copyrightyear{-} 
%dep_docclass\copyrightdata{[to be supplied]}

@@

%\authorinfo{~}{~}{~}
%dep_docclass\authorinfo{Nathan Fulton\and Cyrus Omar\and Jonathan Aldrich}
%dep_docclass           {School of Computer Science\\
%dep_docclass            Carnegie Mellon University}
%dep_docclass           {\{nathanfu, comar, aldrich\}@cs.cmu.edu}

@@

there's extra space here because of the next page...

@@

% I can't achieve the correct latout here. We should have something like:
%       Lam_S Synax         Lam_P Syntax
%       Lam_S Sematics      LAm_P Semantics
%                Translation rules
% Or, alternatively, we can flip reflect things so that the entire page
% reads top-down (Lam_S syntax Lam_S smantics horizontally aligned w/ one
% another).

% Unfortunately, I can't get that to work. So, we're settling for something
% slightly ugly, where the fitures appear in the incorrect order and the 
% translation rules are on top.

% The translations rules *just won't* go to the bottom of the page, so 
% instead I am placing them at the top. Since the figures are out of order,
% it's even more important that we get the number gith. therefore, the counter
% is modified before and after this figure.
\addtocounter{figure}{4}
\begin{figure*}

\end{mathpar}
\caption{Translation from source terms (S) to target terms (P).
The translation is type-directed in the Tr-Coerce cases.}
\end{figure*}
\addtocounter{figure}{-5}

@@

@@

\inferrule[P-T-Check]
{\tctx \vdash P_1 : \regex \\ \tctx \vdash P_2 : \str \\ }
{\tctx \vdash \tcheck{P_1}{P_2} : \str}
\end{mathpar}
\caption{Typing rules for our fragment of $\lambdap$.
The typing context $\tctx$ is standard.}
\label{fig:etyping}
\end{figure}

@@

\newpage
\ 
\newpage
\ 
\newpage
Note that we never insert an internal check where the type of the string implies
that a check must succeed. Furthermore, expensive calculations (such as language
inclusion or ${\tt dsubst}$, all occur at compile time.
Since sanitation problems are generally expressible with smalle, simple expressions,
we believe that the compile-time overhead is not significant enough to prohibiour target usecase. Informal experimentation with our implementation seems to
support this assertion.

@@

\begin{thm}
  Let $S$ be a term in the source language. If $\sctx \vdash S : \stringin{r}$ 
  then $S \sreduces \rstr{s}$ and $\rstr{s} : \stringin{r}$, or else $S \sreduces \err$.
\end{thm}
\begin{proof}
By induction on the derivation of $\sctx \vdash S : \psi$. The interesting case
is S-T-Replace, which requires Lemma C.

\begin{description}
\item[S-T-Stringin-I:]
If $S = \rsconcat{S_1}{S_2} : \stringin{r}$ then
$S \treduces S$ by S-E-RStr, and $\sctx \vdash S : \psi$ by assumption.

\item[S-T-Concat:] 
Suppose $S = \rsconcat{S_1}{S_2} : \stringin{r_1 r_2}$.
By inversion, $\sctx \vdash S_1 : \stringin{r_1}$ and $\sctx \vdash S_1 : \stringin{r_2}$.
It follows by induction that either $S_1 \treduces \err$, $S_2 \treduces \err$, 
or $S_1 \treduces \rstr{s_1}$ and $S_2 \treduces \rstr{s_2}$ for some $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$.
In the latter case $S \reduces \rstr{s_1s_2}$ by S-E-Concat and $\sctx \vdash \rstr{s_1s_2} : \tstr{r_1r_2}$ by Lemma B.
In the former cases, $S \sreduces \err$.

\item[S-T-Replace]:
Suppose $S = \rreplace{r}{S_1}{S_2}$ and $\sctx \vdash S : \stringin{r'}$.
By inversion $\sctx \vdash S_1 : \stringin{r_1}$ and $\sctx \vdash S_2 : \stringin{r_2}$ such that $\lsubst{r}{r_1}{r_2} = r'$.
By induction, $S_1 \sreduces \err$, $S_2 \sreduces \err$ 
or $S_1 \sreduces \rstr{s_1}$ and $S_2 \sreduces \rstr{s_2}$ such that 
In the latter case, we know $\metareplace{r}{s_1}{s_2} \in \lang{\lsubst{r}{r_1}{r_2}}$ by Lemma C;
therefore by S-E-Replace, $S \treduces \rstr{s}$ such that  $s \in \lang{\lsubst{r}{r_1}{r_2}} = \lang{r'}$.
So by S-T-String-I, $\rstr{s} : \stringin{r'}$.
In the former cases, $S \treduces \err$.

\item[S-T-Coerce]:
Suppose $S = \coerce{r}{S_1}$ and $S : \stringin{r}$.
By inversion, $\sctx \vdash S_1 : \stringin{r'}$.
By induction, $S_1 \treduces \err$ or $S_1 \treduces \rstr{s}$.
In the former case $S \treduces \err$ by propagation rules.
In the latter case we have by Lemma B(2) that $s \in \lang{r}$ or else
$s \not \in \lang{r}$.
If $s \in \lang{r}$ then $\rstr{s} : \stringin{r}$.
If $s \not \in \lang{r}$ then $S \treduces \err$. 
\end{description}
\end{proof}

\begin{thm}
  Let $P$ be a term in the target language. If $\tctx \vdash P : \theta$ 
  then $P \sreduces P'$ and $P' : \theta$.
\end{thm}
\begin{proof}
The proof proceeds by induction on the typing relation and is trivial give
and inversion lemma for the typing relation. \texbf{We can write up this proof if we end up having enough space...}
\end{proof}
